---
title: "Feature Extraction"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## degreedata.xlsx    


```{r degree}
library(tidyverse)
library(readxl)
p2_degree = read_excel(path = "feature/p2-fro_degreedata.xlsx", col_names = TRUE)
p3_degree = read_excel(path = "feature/p3-fro_degreedata.xlsx", col_names = TRUE)
p4_degree = read_excel(path = "feature/p4-fro_degreedata.xlsx", col_names = TRUE)
p5_degree = read_excel(path = "feature/p5-fro_degreedata.xlsx", col_names = TRUE)
p6_degree = read_excel(path = "feature/p6-fro_degreedata.xlsx", col_names = TRUE)
p7_degree = read_excel(path = "feature/p7-fro_degreedata.xlsx", col_names = TRUE)
```


### Count of branchpoints     

```{r}
# Count of branchpoints
nrow(p2_degree)
nrow(p3_degree)
nrow(p4_degree)
nrow(p5_degree)
nrow(p6_degree)
nrow(p7_degree)
```

### Spatial density of nodes (Branching number per node)

```{r}
# Branching number per node 
# Spatial density of nodes
deg_spatial_density = data.frame(miceage = c(), degree = c(), density = c())

degree = p2_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("P2",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p3_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("P3",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p4_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("P4",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p5_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("P5",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p6_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("P6",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p7_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("P7",max(degree)), degree = 1:max(degree), density = deg_density))

deg_spatial_density$miceage = as.factor(deg_spatial_density$miceage)
deg_spatial_density$degree = as.factor(deg_spatial_density$degree)
ggplot(deg_spatial_density, aes(x = degree, y = density, color = miceage, group = miceage)) +
  geom_line() +
  geom_point() +
  labs(x = "Number k of vessels branching out for one node",
       y = "% of branching points with node density >= k") +
  scale_colour_discrete("Mice age")+scale_color_manual(values=c("red","orange","magenta","green","blue","black"))
```

Reference: https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=6076630


## alldata.xlsx    


```{r}
p2_all = read_excel(path = "feature/p2-fro_alldata.xlsx", col_names = TRUE)
p3_all = read_excel(path = "feature/p3-fro_alldata.xlsx", col_names = TRUE)
p4_all = read_excel(path = "feature/p4-fro_alldata.xlsx", col_names = TRUE)
p5_all = read_excel(path = "feature/p5-fro_alldata.xlsx", col_names = TRUE)
p6_all = read_excel(path = "feature/p6-fro_alldata.xlsx", col_names = TRUE)
p7_all = read_excel(path = "feature/p7-fro_alldata.xlsx", col_names = TRUE)
```

### Vessel segment - the number of vessel edges

```{r}
nrow(p2_all)
nrow(p3_all)
nrow(p4_all)
nrow(p5_all)
nrow(p6_all)
nrow(p7_all)
```

It certainly increases as it gets older.        

### Vessel length & Tortuosity & Width

```{r}
mean(p2_all$length)
mean(p3_all$length)
mean(p4_all$length)
mean(p5_all$length)
mean(p6_all$length)
mean(p7_all$length)

sum(p2_all$length)
sum(p3_all$length)
sum(p4_all$length)
sum(p5_all$length)
sum(p6_all$length)
sum(p7_all$length)

mean(p2_all$tortuosity)
mean(p3_all$tortuosity)
mean(p4_all$tortuosity)
mean(p5_all$tortuosity)
mean(p6_all$tortuosity)
mean(p7_all$tortuosity)

mean(p2_all$width)
mean(p3_all$width)
mean(p4_all$width)
mean(p5_all$width)
mean(p6_all$width)
mean(p7_all$width)
```

Sum of vessel length increases as it gets older   
Couldn't find any other interesting facts (differences) for length and tortuosity.    
Maybe Vessel width (diameter) gets smaller as it gets older? (perform t-test?)    


## Summary 

```{r}
vessel_summary = data.frame(p2 = c(),p3 = c(),p4 = c(),p5 = c(),p6 = c(),p7 = c())
f1 = c(nrow(p2_degree),nrow(p3_degree),nrow(p4_degree),nrow(p5_degree),nrow(p6_degree),nrow(p7_degree))
f2 = c(nrow(p2_all),nrow(p3_all),nrow(p4_all),nrow(p5_all),nrow(p6_all),nrow(p7_all))
f3 = c(mean(p2_all$length),mean(p3_all$length),mean(p4_all$length),
       mean(p5_all$length),mean(p6_all$length),mean(p7_all$length))
f4 = c(mean(p2_all$tortuosity),mean(p3_all$tortuosity),mean(p4_all$tortuosity),
       mean(p5_all$tortuosity),mean(p6_all$tortuosity),mean(p7_all$tortuosity))
f5 = c(mean(p2_all$width),mean(p3_all$width),mean(p4_all$width),
       mean(p5_all$width),mean(p6_all$width),mean(p7_all$width))
  
  
m = rbind(f1,f2,f3,f4,f5)
colnames(m) = c("P2","P3","P4","P5","P6","P7")
rownames(m) = c("Number of Branchpoints", "Number of Vessel Edges","Mean of Vessel Length",
                "Mean of Vessel Tortuosity","Mean of Vessel Width")

options(digits=4)
knitr::kable(as.data.frame(m), align="c")

```   

## Density and Boxplot

```{r}
p1to7_all <- rbind(p2_all, p3_all, p4_all, p5_all, p6_all, p7_all)
p1to7_all$stage <- c(rep("P2", nrow(p2_all)), rep("P3", nrow(p3_all)), rep("P4", nrow(p4_all)), rep("P5", nrow(p5_all)), rep("P6", nrow(p6_all)), rep("P7", nrow(p7_all)))

colnames(p1to7_all) <- c("nodespair", "node1", "node2", "Vessel.Line", "Vessel.Length", "Vessel.Width", "wide_var", "Vessel.Tortuosity", "curve", "Mice.age")

p1to7_all_long <- gather(p1to7_all, key = "feature", value = "value",
                         Vessel.Line, Vessel.Length, Vessel.Width, Vessel.Tortuosity)
ggplot(p1to7_all_long, mapping = aes(x = log(value), fill = Mice.age, color = Mice.age)) + facet_wrap(.~feature, nrow = 2, scales = "free")  + geom_line(stat = "density")+theme(axis.title.x = element_blank(), legend.position = "bottom")+ guides(color=guide_legend(nrow=1,byrow=TRUE))+scale_color_manual(name="Mice age",values=c("red","orange","magenta","green","blue","black"))+ ylab("Density of log(value)")

ggplot(p1to7_all_long, aes(x=Mice.age, y=log(value), fill=Mice.age)) + facet_wrap(.~feature, nrow = 2, scales = "free") + geom_boxplot(width = 0.8, outlier.size=0.4) +theme(axis.title.x = element_blank(),legend.position = "bottom")+  scale_x_discrete(expand=c(0.12, 0))+ scale_fill_manual(name="Mice age",values = c("red","orange","magenta","green","blue","black"))+ guides(fill=guide_legend(nrow=1,byrow=TRUE))
```

It seems that width variable has less variance as mouse grows.


## Area

```{r}
## Install EBImage package
### http://bioconductor.org/packages/release/bioc/html/EBImage.html
library(EBImage)
my_image2 <- readImage("../data/test_data/predictions/p2-from 5-5-2 M_prediction.tif")
my_image3 <- readImage("../data/test_data/predictions/p3-from 5-5-2 M_prediction.tif")
my_image4 <- readImage("../data/test_data/predictions/p4-from 5-5-2 M_prediction.tif")
my_image5 <- readImage("../data/test_data/predictions/p5-from 5-5-1 M_prediction.tif")
my_image6 <- readImage("../data/test_data/predictions/p6-from 5-5-2 M_prediction.tif")
my_image7 <- readImage("../data/test_data/predictions/p7-from 5-5-2 M_prediction.tif")

options(digits=4)
knitr::kable(data.frame(p2 = mean(my_image2), p3 = mean(my_image3), p4 = mean(my_image4),
                        p5 = mean(my_image5), p6 = mean(my_image6), p7 = mean(my_image7)), align = "c")
```

VAD: ratio of the total image area occupied by blood vessel area
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4901200/pdf/JBO-021-066008.pdf (page 4)



## Bar plots

```{r}
library(gridExtra)
library(cowplot)

## VAD: ratio of the total image area occupied by blood vessel area
f6 = c(mean(my_image2), mean(my_image3), mean(my_image4),
       mean(my_image5), mean(my_image6), mean(my_image7))

## 
## 
# f7 = c(max(p2_degree$distance), max(p3_degree$distance), max(p4_degree$distance), max(p5_degree$distance), max(p6_degree$distance), max(p7_degree$distance))


f7 = c(quantile(p2_degree$distance, 0.90), quantile(p3_degree$distance, 0.90), max(p4_degree$distance), 
       max(p5_degree$distance), max(p6_degree$distance), max(p7_degree$distance))
      
# f7 = c(quantile(p2_degree$distance, 0.95), quantile(p3_degree$distance, 0.95), quantile(p4_degree$distance, 0.95), quantile(p5_degree$distance, 0.95), quantile(p6_degree$distance, 0.95), quantile(p7_degree$distance, 0.95))

df <- data.frame(stage=c("P2","P3","P4","P5","P6","P7"),
                 f1 = f1, f2 = f2, f3 = f3, f4 = f4, f5 = f5, f6 = f6, f7 = f7)
colnames(df) <- c("stage", "Number.of.Branchpoints", "Number.of.Vessel.Edges", "Mean.of.Vessel.Length",
                  "Mean.of.Vessel.Tortuosity", "Mean.of.Vessel.Width", "VAD", "Distance.from.center.to.end")
df_long <- gather(df, key = "feature", value = "value",
                  Number.of.Branchpoints, Number.of.Vessel.Edges, Mean.of.Vessel.Length,
                  Mean.of.Vessel.Tortuosity, Mean.of.Vessel.Width, VAD, Distance.from.center.to.end)

df_long$feature <- factor(df_long$feature, levels = c("Number.of.Branchpoints",              
                                                     "Number.of.Vessel.Edges", 
                                                     "Mean.of.Vessel.Length", 
                                                     "Mean.of.Vessel.Tortuosity", 
                                                     "Mean.of.Vessel.Width", 
                                                     "VAD",  
                                                     "Distance.from.center.to.end"))

feature.p <- ggplot(data=df_long, aes(x=stage, y=value, fill = stage)) +
  geom_bar(stat="identity", width=0.8, show.legend=FALSE)  + facet_wrap(.~feature, nrow = 1, scales = "free")+theme(axis.title.x = element_blank(), strip.text = element_text(size=7.5), panel.spacing.y = unit(1, "lines")) + scale_fill_manual(name="Mice age", values = c("red","orange","magenta","green","blue","black"))


library(egg)
library(gridExtra)
feature.p1 <- feature.p  %+% df_long[1:24,] + labs(x = NULL)
feature.p2 <- feature.p  %+% df_long[-(1:24),]
grid.arrange(grobs = lapply(
  list(feature.p1, feature.p2),
  set_panel_size,
  width = unit(3.6, "cm"),
  height = unit(4, "cm")
))

df.rate <- data.frame(stage=c("P2","P3","P4","P5","P6","P7"),
                 f1 = c(0,f1[2:6]/f1[1:5]-1), f2 = c(0,f2[2:6]/f2[1:5]-1), 
                 f3 = c(0,f3[2:6]/f3[1:5]-1), f4 = c(0,f4[2:6]/f4[1:5]-1), 
                 f5 = c(0,f5[2:6]/f5[1:5]-1), f6 = c(0,f6[2:6]/f6[1:5]-1), 
                 f7 = c(0,f7[2:6]/f7[1:5]-1))

colnames(df.rate) <- c("stage", "Number.of.Branchpoints", "Number.of.Vessel.Edges", "Mean.of.Vessel.Length",
                  "Mean.of.Vessel.Tortuosity", "Mean.of.Vessel.Width", "VAD", "Distance.from.center.to.end")

df.rate_long <- gather(df.rate, key = "feature", value = "Increase.rate",
                  Number.of.Branchpoints, Number.of.Vessel.Edges, Mean.of.Vessel.Length,
                  Mean.of.Vessel.Tortuosity, Mean.of.Vessel.Width, VAD, Distance.from.center.to.end)

df.rate_long$feature <- factor(df.rate_long$feature, levels = c("Number.of.Branchpoints",              
                                                     "Number.of.Vessel.Edges", 
                                                     "Mean.of.Vessel.Length", 
                                                     "Mean.of.Vessel.Tortuosity", 
                                                     "Mean.of.Vessel.Width", 
                                                     "VAD",  
                                                     "Distance.from.center.to.end"))

rate.p <- ggplot(data=df.rate_long, aes(x=stage, y=Increase.rate, group = 1)) +
  geom_point(aes(colour = factor(stage))) + geom_line() + facet_wrap(.~feature, nrow = 1) + theme(axis.title.x = element_blank(), strip.text = element_text(size=7.5)) + scale_color_manual(name="Mice age", values = c("red","orange","magenta","green","blue","black"))+ scale_y_continuous(limits = c(-0.5, 1.5)) + theme(legend.position = "none",panel.spacing.y = unit(1, "lines"))


library(egg)
library(gridExtra)
rate.p1 <- rate.p  %+% df.rate_long[1:24,] + labs(x = NULL)
rate.p2 <- rate.p  %+% df.rate_long[-(1:24),]
grid.arrange(grobs = lapply(
  list(rate.p1, rate.p2),
  set_panel_size,
  width = unit(3.6, "cm"),
  height = unit(4, "cm")
))

```   

=======
<<<<<<< HEAD
=======
