---
title: "Feature Extraction"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## degreedata.xlsx    


```{r degree}
library(tidyverse)
library(readxl)
p2_degree = read_excel(path = "feature/p2-fro_degreedata.xlsx", col_names = TRUE)
p3_degree = read_excel(path = "feature/p3-fro_degreedata.xlsx", col_names = TRUE)
p4_degree = read_excel(path = "feature/p4-fro_degreedata.xlsx", col_names = TRUE)
p5_degree = read_excel(path = "feature/p5-fro_degreedata.xlsx", col_names = TRUE)
p6_degree = read_excel(path = "feature/p6-fro_degreedata.xlsx", col_names = TRUE)
p7_degree = read_excel(path = "feature/p7-fro_degreedata.xlsx", col_names = TRUE)
```


### Count of branchpoints     

```{r}
# Count of branchpoints
nrow(p2_degree)
nrow(p3_degree)
nrow(p4_degree)
nrow(p5_degree)
nrow(p6_degree)
nrow(p7_degree)
```

### Spatial density of nodes (Branching number per node)

```{r}
# Branching number per node 
# Spatial density of nodes
deg_spatial_density = data.frame(miceage = c(), degree = c(), density = c())

degree = p2_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("p2",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p3_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("p3",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p4_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("p4",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p5_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("p5",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p6_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("p6",max(degree)), degree = 1:max(degree), density = deg_density))

degree = p7_degree$degree
deg_density = table(degree)[max(degree):1] %>%
  prop.table() %>%
  cumsum() %>%
  rev()
deg_spatial_density = rbind(deg_spatial_density, 
                            data.frame(miceage = rep("p7",max(degree)), degree = 1:max(degree), density = deg_density))

deg_spatial_density$miceage = as.factor(deg_spatial_density$miceage)
deg_spatial_density$degree = as.factor(deg_spatial_density$degree)
ggplot(deg_spatial_density, aes(x = degree, y = density, color = miceage, group = miceage)) +
  geom_line() +
  geom_point() +
  labs(x = "Number k of vessels branching out for one node",
       y = "% of branching points with node density >= k") +
  scale_colour_discrete("Mice age")
```

Reference: https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=6076630


## alldata.xlsx    


```{r}
p2_all = read_excel(path = "feature/p2-fro_alldata.xlsx", col_names = TRUE)
p3_all = read_excel(path = "feature/p3-fro_alldata.xlsx", col_names = TRUE)
p4_all = read_excel(path = "feature/p4-fro_alldata.xlsx", col_names = TRUE)
p5_all = read_excel(path = "feature/p5-fro_alldata.xlsx", col_names = TRUE)
p6_all = read_excel(path = "feature/p6-fro_alldata.xlsx", col_names = TRUE)
p7_all = read_excel(path = "feature/p7-fro_alldata.xlsx", col_names = TRUE)
```

### Vessel segment - the number of vessel edges

```{r}
nrow(p2_all)
nrow(p3_all)
nrow(p4_all)
nrow(p5_all)
nrow(p6_all)
nrow(p7_all)
```

It certainly increases as it gets older.        

### Vessel length & Tortuosity & Width

```{r}
mean(p2_all$length)
mean(p3_all$length)
mean(p4_all$length)
mean(p5_all$length)
mean(p6_all$length)
mean(p7_all$length)

sum(p2_all$length)
sum(p3_all$length)
sum(p4_all$length)
sum(p5_all$length)
sum(p6_all$length)
sum(p7_all$length)

mean(p2_all$tortuosity)
mean(p3_all$tortuosity)
mean(p4_all$tortuosity)
mean(p5_all$tortuosity)
mean(p6_all$tortuosity)
mean(p7_all$tortuosity)

mean(p2_all$width)
mean(p3_all$width)
mean(p4_all$width)
mean(p5_all$width)
mean(p6_all$width)
mean(p7_all$width)
```

Sum of vessel length increases as it gets older   
Couldn't find any other interesting facts (differences) for length and tortuosity.    
Maybe Vessel width (diameter) gets smaller as it gets older? (perform t-test?)    


## Summary 

```{r}
vessel_summary = data.frame(p2 = c(),p3 = c(),p4 = c(),p5 = c(),p6 = c(),p7 = c())
f1 = c(nrow(p2_degree),nrow(p3_degree),nrow(p4_degree),nrow(p5_degree),nrow(p6_degree),nrow(p7_degree))
f2 = c(nrow(p2_all),nrow(p3_all),nrow(p4_all),nrow(p5_all),nrow(p6_all),nrow(p7_all))
f3 = c(mean(p2_all$length),mean(p3_all$length),mean(p4_all$length),
       mean(p5_all$length),mean(p6_all$length),mean(p7_all$length))
f4 = c(mean(p2_all$tortuosity),mean(p3_all$tortuosity),mean(p4_all$tortuosity),
       mean(p5_all$tortuosity),mean(p6_all$tortuosity),mean(p7_all$tortuosity))
f5 = c(mean(p2_all$width),mean(p3_all$width),mean(p4_all$width),
       mean(p5_all$width),mean(p6_all$width),mean(p7_all$width))
  
  
m = rbind(f1,f2,f3,f4,f5)
colnames(m) = c("p2","p3","p4","p5","p6","p7")
rownames(m) = c("Number of Branchpoints", "Number of Vessel Edges","Mean of Vessel Length",
                "Mean of Vessel Tortuosity","Mean of Vessel Width")

options(digits=4)
knitr::kable(as.data.frame(m), align="c")

```   

## Density and Boxplot

```{r}
p1to7_all <- rbind(p2_all, p3_all, p4_all, p5_all, p6_all, p7_all)
p1to7_all$stage <- c(rep("p2", nrow(p2_all)), rep("p3", nrow(p3_all)), rep("p4", nrow(p4_all)), rep("p5", nrow(p5_all)), rep("p6", nrow(p6_all)), rep("p7", nrow(p7_all)))
p1to7_all_long <- gather(p1to7_all, key = "feature", value = "value",
       line, length, width, tortuosity)
ggplot(p1to7_all_long, mapping = aes(x = value, fill = stage, color = stage)) + facet_wrap(.~feature, nrow = 2, scales = "free")  + geom_line(stat = "density")+theme(legend.position = "bottom")+ guides(fill=guide_legend(nrow=1,byrow=TRUE))
ggplot(p1to7_all_long, aes(x=stage, y=value, fill=stage)) + facet_wrap(.~feature, nrow = 2, scales = "free") +
  geom_boxplot(width = 0.8, outlier.size=0.4) +theme(legend.position = "bottom")+  scale_x_discrete(expand=c(0.24, 0)) + scale_fill_manual(values = c("#FC4E07", "#E69F00", "#F0E442", "#009E73", "#56B4E9","#CC79A7")) + guides(fill=guide_legend(nrow=1,byrow=TRUE))
```

It seems that width variable has less variance as mouse grows.


## Area

```{r}
## Install EBImage package
### http://bioconductor.org/packages/release/bioc/html/EBImage.html
library(EBImage)
my_image2 <- readImage("../data/test_data/predictions/p2-from 5-5-2 M_prediction.tif")
my_image3 <- readImage("../data/test_data/predictions/p3-from 5-5-2 M_prediction.tif")
my_image4 <- readImage("../data/test_data/predictions/p4-from 5-5-2 M_prediction.tif")
my_image5 <- readImage("../data/test_data/predictions/p5-from 5-5-1 M_prediction.tif")
my_image6 <- readImage("../data/test_data/predictions/p6-from 5-5-2 M_prediction.tif")
my_image7 <- readImage("../data/test_data/predictions/p7-from 5-5-2 M_prediction.tif")

options(digits=4)
knitr::kable(data.frame(p2 = mean(my_image2), p3 = mean(my_image3), p4 = mean(my_image4),
                        p5 = mean(my_image5), p6 = mean(my_image6), p7 = mean(my_image7)), align = "c")
```

VAD: ratio of the total image area occupied by blood vessel area
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4901200/pdf/JBO-021-066008.pdf (page 4)



## Bar plots

```{r}
library(gridExtra)
library(cowplot)

## VAD: ratio of the total image area occupied by blood vessel area
f6 = c(mean(my_image2), mean(my_image3), mean(my_image4),
       mean(my_image5), mean(my_image6), mean(my_image7))

## 
f7 = c(max(p2_degree$distance), max(p3_degree$distance), max(p4_degree$distance), 
       max(p5_degree$distance), max(p6_degree$distance), max(p7_degree$distance))

df <- data.frame(miceage=c("p2","p3","p4","p5","p6","p7"),
                 f1 = f1, f2 = f2, f3 = f3, f4 = f4, f5 = f5, f6 = f6, f7 = f7)


p1 <- ggplot(data=df, aes(x=miceage, y=f1, fill = miceage)) +
  geom_bar(stat="identity", width=0.8, show.legend=FALSE) +
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Mice Age", y = "Number of Branchpoints") 

p2 <- ggplot(data=df, aes(x=miceage, y=f2, fill = miceage)) +
  geom_bar(stat="identity", width=0.8, show.legend=FALSE) +
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Mice Age", y = "Number of Vessel Edges")

p3 <- ggplot(data=df, aes(x=miceage, y=f3, fill = miceage)) +
  geom_bar(stat="identity", width=0.8, show.legend=FALSE) +
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Mice Age", y = "Mean of Vessel Length")

p4 <- ggplot(data=df, aes(x=miceage, y=f4, fill = miceage)) +
  geom_bar(stat="identity", width=0.8, show.legend=FALSE) +
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Mice Age", y = "Mean of Vessel Tortuosity")

p5 <- ggplot(data=df, aes(x=miceage, y=f5, fill = miceage)) +
  geom_bar(stat="identity", width=0.8, show.legend=FALSE) +
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Mice Age", y = "Mean of Vessel Width")
  
p6 <- ggplot(data=df, aes(x=miceage, y=f6, fill = miceage)) +
  geom_bar(stat="identity", width=0.8, show.legend=FALSE) +
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Mice Age", y = "VAD")

p7 <- ggplot(data=df, aes(x=miceage, y=f7, fill = miceage)) +
  geom_bar(stat="identity", width=0.8, show.legend=FALSE) +
  scale_fill_brewer(palette="Blues") + 
  labs(x = "Mice Age", y = "Distance from center to end")

#grid.arrange(p1,p2,p3,p4,p5,p6,p7, ncol = 1)
#plot_grid(p1,p2,p3,p4,p5,p6,p7, nrow = 4, labels = "AUTO")

p1
p2
p3
p4
p5
p6
p7

ggplot(data=df, aes(x=c("p2","p3","p4","p5","p6","p7"), y=c(0,f1[2:6]/f1[1:5]-1), group = 1)) +
  geom_point() +
  geom_line() +
  labs(x = "Mice Age", y = "Increase rate") +
  ylim(-0.5,1.5) 

ggplot(data=df, aes(x=c("p2","p3","p4","p5","p6","p7"), y=c(0,f2[2:6]/f2[1:5]-1), group = 1)) +
  geom_point() +
  geom_line() +
  labs(x = "Mice Age", y = "Increase rate") +
  ylim(-0.5,1.5) 

ggplot(data=df, aes(x=c("p2","p3","p4","p5","p6","p7"), y=c(0,f3[2:6]/f3[1:5]-1), group = 1)) +
  geom_point() +
  geom_line() +
  labs(x = "Mice Age", y = "Increase rate") +
  ylim(-0.5,1.5) 

ggplot(data=df, aes(x=c("p2","p3","p4","p5","p6","p7"), y=c(0,f4[2:6]/f4[1:5]-1), group = 1)) +
  geom_point() +
  geom_line() +
  labs(x = "Mice Age", y = "Increase rate") +
  ylim(-0.5,1.5) 

ggplot(data=df, aes(x=c("p2","p3","p4","p5","p6","p7"), y=c(0,f5[2:6]/f5[1:5]-1), group = 1)) +
  geom_point() +
  geom_line() +
  labs(x = "Mice Age", y = "Increase rate") +
  ylim(-0.5,1.5) 

ggplot(data=df, aes(x=c("p2","p3","p4","p5","p6","p7"), y=c(0,f6[2:6]/f6[1:5]-1), group = 1)) +
  geom_point() +
  geom_line() +
  labs(x = "Mice Age", y = "Increase rate") +
  ylim(-0.5,1.5)

ggplot(data=df, aes(x=c("p2","p3","p4","p5","p6","p7"), y=c(0,f7[2:6]/f7[1:5]-1), group = 1)) +
  geom_point() +
  geom_line() +
  labs(x = "Mice Age", y = "Increase rate") +
  ylim(-0.5,1.5)
  

```   

=======
>>>>>>> dd966e4653d2f0f66ce62b7246cceebe2b712041
